SHARP - II. Mass structure in strong lenses is not necessarily dark matter substructure: A flux ratio anomaly from an edge-on disc in B1555+375

### 2016-06-22 papers/B1555_EdgeOn/B1555_letter.tex
change legend setting in plot_fns_b1555.py
minor text change (shrink) in B1555 paper

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Beware the effect of edge-on discs]{SHARP - II. Mass structure in strong lenses is not necessarily dark matter substructure: A flux ratio anomaly from an edge-on disc }
\author[Hsueh et al.]{J.-W. Hsueh,$^{1}$\thanks{E-mail: jwhsueh@ucdavis.edu} C. D. Fassnacht,$^{1}$ S. Vegetti,$^{2}$  J. P. McKean,$^{3,4}$ C. Spingola,$^{4}$ M. W. Auger,$^{5}$
\newauthor  L. V. E. Koopmans,$^{4}$ and D. J. Lagattuta$^{6}$\\
$^{1}$Department of Physics, University of California, Davis, 1 Shields Ave. Davis, CA 95616, USA\\
$^{2}$Max Planck Institute for Astrophysics, Karl-Schwarzschild-Strasse 1, D-85740 Garching, Germany\\
$^{3}$Netherlands Institute for Radio Astronomy (ASTRON), P.O. Box 2, 7990 AA Dwingeloo, The Netherlands\\
$^{4}$Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands\\
$^{5}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{6}$University of Lyon, CRAL, Observatoire de Lyon, 92 Rue Pasteur, 69007 Lyon, France}
\begin{document}

@@

The next step is to test if a physically plausible edge-on disc can cause a perturbation in the strong lensing mass model that produces the observed flux ratio anomaly of  merging images A and B. An exponential disc profile best describes the disc component in spiral galaxies because it matches the light distribution \citep{Kee98}. We thus choose a SIE plus an exponential disc model (SIE+ExpDisc) in our study of B1555+375. The free parameters of this model are the SIE Einstein radius ($b$), centroid position, ellipticity ($e = 1 - q$), position angle ($\theta$), the exponential disc intrinsic central density ($\kappa_0$) and scale length ($R_s$). A detailed definition of these mass profiles is described by \citet{Kee01}. The source position adds two additional model parameters. We have not included a third component such as an external shear or additional NFW halo term because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model.
%An SIE profile is also sufficient to include the halo mass since B1555 has a small Einstein radius.}

More sophisticated mass models require additional
constraints, such as those from the lensed extended emission in either
the NIR or radio imaging.  While we have successfully used the former
approach in the past \citep[e.g.][]{V12}, for B1555+375 the extinction
in the AO imaging is so severe that components B and D are not
visible, and the arc that connects A and B appears to be truncated.
The low signal-to-noise ratio in the ring away from the lensed AGN and
the strong extinction in the lens galaxy prevent us from using this
gravitational imaging approach.  Significantly deeper
observations are needed at a higher bandwidth with very long baseline
interferometric arrays, which can test more complex models in the
future by imaging the extended gravitational arc and more
significantly detecting image D.

\textbf{We use {\sc gravlens} to obtain the best-fit SIE+ExpDisc lens model. The best-fit parameters from \citet{Marlow99} are chosen as the starting point of the SIE profile. We have used the NICMOS F160W imaging, in which the disc component has the highest signal-to-noise ratio amongst our data, to fix the disc centroid position relative to image A to $(-0.162, -0.206)$, its ellipticity to $e=0.83$, and its position angle to $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain the uncertainties.}

The best-fit model parameters and 68 per cent confidence level (CL) errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in the fluxes may be due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the \textbf{ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$ but the mass fraction is independent of the lens and source redshifts since this ratio is independent of distance).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. Investigations of edge-on spiral lenses, for example from the SWELLS survey, show that the disc mass fraction within the Einstein radius ranges from 5 to 30 per cent\citep{swells1,swells3,swells6}.} 
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the MERLIN radio position of image A. The ellipticity is defined as $e=1-q$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius ($R_s$) are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.177 \pm 0.003$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.205 \pm 0.004$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.178 \pm 0.004$  		& ...  \\
$\kappa_0$ & ... &$0.273 \pm 0.003$\\
$e$	  		& $0.25 \pm 0.05$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.9 \pm 1.3$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.236 \pm 0.003$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Beware the effect of edge-on discs]{SHARP - II. Mass structure in strong lenses is not necessarily dark matter substructure: A flux ratio anomaly from an edge-on disc in B1555+375}
\author[Hsueh et al.]{J.-W. Hsueh,$^{1}$\thanks{E-mail: jwhsueh@ucdavis.edu} C. D. Fassnacht,$^{1}$ S. Vegetti,$^{2}$  J. P. McKean,$^{3,4}$ C. Spingola,$^{4}$ M. W. Auger,$^{5}$
\newauthor  L. V. E. Koopmans,$^{4}$ and D. J. Lagattuta$^{6}$\\
$^{1}$Department of Physics, University of California, Davis, 1 Shields Ave. Davis, CA 95616, USA\\
$^{2}$Max Planck Institute for Astrophysics, Karl-Schwarzschild-Strasse 1, D-85740 Garching, Germany\\
$^{3}$Netherlands Institute for Radio Astronomy (ASTRON), P.O. Box 2, 7990 AA Dwingeloo, The Netherlands\\
$^{4}$Kapteyn Astronomical Institute, University of Groningen, P.O. Box 800, 9700 AV Groningen, The Netherlands\\
$^{5}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{6}$University of Lyon, CRAL, Observatoire de Lyon, 92 Rue Pasteur, 69007 Lyon, France}
\begin{document}

@@

The next step is to test if a physically plausible edge-on disc can cause a perturbation in the strong lensing mass model that produces the observed flux ratio anomaly of  merging images A and B. \textbf{Our model contains a SIE plus an exponential disc component (SIE+ExpDisc) due to that  exponential disc profile matches the light distribution of the disc component in spiral galaxies \citep{Kee98}.} The free parameters of this model are the SIE Einstein radius ($b$), centroid position, ellipticity ($e = 1 - q$), position angle ($\theta$), the exponential disc intrinsic central density ($\kappa_0$) and scale length ($R_s$). A detailed definition of these mass profiles is described by \citet{Kee01}. The source position adds two additional model parameters. 
%We have not included a third component such as an external shear or additional NFW halo term because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model.
%An SIE profile is also sufficient to include the halo mass since B1555 has a small Einstein radius.}

More sophisticated mass models \textbf{(i.e. external shear or additional halo term)} require additional
constraints, such as those from the lensed extended emission in either
the NIR or radio imaging.  While we have successfully used the former
approach in the past \citep[e.g.][]{V12}, for B1555+375 the extinction
in the AO imaging is so severe that components B and D are not
visible, and the arc that connects A and B appears to be truncated.
The low signal-to-noise ratio in the ring away from the lensed AGN and
the strong extinction in the lens galaxy prevent us from using this
gravitational imaging approach.  Significantly deeper
observations are needed at a higher bandwidth with very long baseline
interferometric arrays, which can test more complex models in the
future by imaging the extended gravitational arc and more
significantly detecting image D.

We use {\sc gravlens} to obtain the best-fit SIE+ExpDisc lens model. The best-fit parameters from \citet{Marlow99} are chosen as the starting point of the SIE profile. We have used the NICMOS F160W imaging, in which the disc component has the highest signal-to-noise ratio amongst our data, to fix the disc centroid position relative to image A to $(-0.162, -0.206)$, its ellipticity to $e=0.83$, and its position angle to $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain the uncertainties.

\textbf{Our model predicted results are shown in Table \ref{tab:results} with comparison to the observed data and the single SIE lens model.}The best-fit model parameters and 68 per cent confidence level (CL) errors are listed in Table~\ref{tab:model}, and the lens model is illustrated in Fig.~\ref{fig:3panel}c.  We find that adding an elongated exponential disc to a smooth SIE model \textbf{successfully reproduces the flux ratio anomaly in this system}. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in the fluxes may be due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the disc component makes up about 15 per cent of the total projected mass within the Einstein radius.
% (1.7~kpc for an assumed lens redshift of $z_l = 1$ but the mass fraction is independent of the lens and source redshifts since this ratio is independent of distance).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. 
\textbf{While} investigations of edge-on spiral lenses, for example from the SWELLS survey, show that the disc mass fraction within the Einstein radius ranges from 5 to 30 per cent\citep{swells1,swells3,swells6}.
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the MERLIN radio position of image A. The position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius ($R_s$) are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.177 \pm 0.003$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.205 \pm 0.004$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.178 \pm 0.004$  		& ...  \\
$\kappa_0$ & ... &$0.273 \pm 0.003$\\
$e$	  		& $0.25 \pm 0.05$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.9 \pm 1.3$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.236 \pm 0.003$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}
